Learning how to act: making good decisions with machine learning

### 2017-09-04 thesis/thesis_map.tex
minor fixes

This thesis is about machine learning and statistical approaches to decision making. How can we learn from data to anticipate the consequence of, and optimally select, interventions or actions? Such problems, from deciding which medication to prescribe to patients, who should be released on bail or how much to charge for insurance, to designing policies to reduce unemployment or improve school attendance are ubiqutes and have far reaching impacts on our lives. There are two fundamental approaches to learning how to act: reinforcment learning, in which an agent directly intervenes in a system and learns from the outcome, and observational causal inference, whereby we seek to infer the outcome of an intervention from observing the system.

The goal of this thesis to to connect and unify these key approaches. I introduce causal bandit problems: a framework which combines causal graphical models, which were developed for observational causal inference, with multi-armed bandit problems, which are a subset of reinforcment learning problems that are simple enough to admit formal analysis. I show that knoweledge of the causal structure allows us to transfer information learned about the outcome of one action to predict the outcome of an alternate action, yielding a novel form of structure between bandit arms that cannot be expoited by existing algorithms. I propose an algorithm for causal bandit problems and prove bounds on the simple regret demonstrating it is close to mini-max optimal and strictly better than algorithms that do not use the additional causal information.

@@

\section{Motivation}
Many of the most important questions in science and in our personal lives are about the outcomes of doing something. Will asking people to pay upfront at the doctors reduce long term health expenditure? If we developed a drug to suppress particular genes, could we cure MS and would delaying teen-aged pregnancies improve the outcome for their kids.

These are hard questions because they require more than identifying a pattern in data. Correlation is not causation. Causal inference has proven so difficult that there is barely any consensus on even enduring questions like the returns to education or the long-term consequences of early life events – like teenage pregnancy, where the variables involved are susceptible to human intuition and understanding.

We now live in a world of data. Hours of our lives are spent online, where every click can be recorded, tiny computers and sensors are cheap enough to incorporate into everything and the US Institute of Health is considering if all infants should be genetically sequenced at birth. Such data gives us a window into many aspects of our lives at an unprecedented scale and detail but it is messy, complicated and often generated as a by product of some other purpose. It does not come from the controlled world of a randomised experiment.

The rise of big data sets and powerful computers has seen an explosion in the application of machine learning. From health care, to entertainment and self driving cars; machine learning algorithms will transform many industries. It has been suggested that the impressive ability of statistical machine learning to detect complex patterns in huge data sets heralds the end of theory \citep{anderson2008end} and that we may be only a short step from \quotes{The Singularity}, where artificial intelligence exceeds our own and then grows exponentially.

However, despite the huge advances in machine learning (in particular deep learning), machine learning algorithms are effective only within narrow problem settings. Getting them to generalise to even slightly different problems or data sets remains very challenging. Deciding how we should act or what policies we should implement requires us to be able to predict how a system will behave if we change it. The correlations detected by standard machine learning algorithms do not enable us to do this, no matter how many petabytes of data they are based on. As machine learning algorithms are incorporated into more and more of the decision making processes that shape the world we live it, it is critical to ensure that we understand the distinction between causality and prediction and that we develop techniques for learning how to act that are as effective as those we have for pattern recognition.

@@

The notion of causality has been widely debated in science and philosophy \citep{hume1741treatise,mill1893system,pearson1911grammar,Pearl2000,Lewis2000,woodward2005making,heckman20051,cartwright2007hunting} but is still widely viewed as poorly defined. This has led to a reluctance among applied researchers in many fields to refer to causality in their work, leading them instead to report that variables are \textit{related}, \textit{correlated} or \textit{associated}. However, the magnitude, direction and even existence of an association depends on what other variables we adjust for (or include in a regression). Avoiding formalising causation, the real question of interest, leaves it up to the reader to determine via \quotes{common sense} the implications of the reported associations.

There are two ways in which an association detected in a data set may be non-causal. The first is that the variables concerned may not be related at all, and the association has arisen by chance in that data sample. Given a finite data on enough variables, there is a high probability of finding some that appear correlated even though they are completely unrelated. For example, based on data from the years 1999 to 2009, the age of miss America is strongly correlated with the number of murders  (in the US) by steam, hot vapours and hot objects \citep{Vigen2015}. We would not expect this relationship to hold if we were to look at a new sample of data. This form of spurious correlation also has serious repercussions. It lies at the heart of major problems with the process of scientific research: researchers are incentivised to detect effects and thus to explore many possible paths in the process of analysing data and studies that fail to find an effect are less likely to be published. As a consequence, the likelyhood that reported effects have arisen by chance is underestimated, leading to the conclusion that \quotes{most published scientific results are false} \citep{Ioannidis2005}. This issue is also highlighted by recent crises in replication \citep{OpenScienceCollaboration2015}. This issue can be ameliorated by getting more data and by separating learning models from evaluating their performance, for example by evaluating models on a strict hold-out set or on the extent to which their results can be replicated.

However, a strong association, observed in multiple independent studies may still not be causal. The correlation can arise because both variables are consequences of some other, unmeasured factor. For example, the reading ability of children under twelve is strongly correlated with their height, because older children are taller and can read better. However height is not a cause of reading ability because if we were to intervene to increase a child's height, for example by giving them growth hormones, we would not expect dramatic improvements in their reading. Similarly, extra tutoring in reading will not make a child grow taller. This problem is fundamentally different to the issue of spurious correlations arising by chance in finite data sets. Obtaining more (even infinitely many more) samples without directly intervening in the system to manipulate the variables does not allow us to separate causation from correlation.

The key distinction between a real, but non-causal, association and a causal relationship is in what happens if we intervene in the system and change one of the variables. In this thesis, I take an interventionist viewpoint of causality: any model or approach designed to predict the outcome of intervening in a system is causal. This viewpoint captures the types of questions that motivate this thesis. How can we change the way we do things to obtain better outcomes.

Causality is often liked to explanation; understanding how and why things happen. I view explanation in terms of compression and generalisabilty: how much information about the world a model can capture. This creates a hierarchy in the degree to which models are explanatory, rather than a simple binary distinction. A standard predictive model encodes all the information we need to predict some output given inputs provided the system generating the data does not change. A high-level causal model might be able to predict the outcome of a specific intervention holding all else fixed. More detailed causal models could predict the outcome for a wide range of combinations of interventions conditional on a range of contexts. By considering conditional interventions within our definition of causal questions we also capture mediation: the study of through which pathways one variable causes another \cite{vanderweele2015explanation}. Finally, a model that can distil how elements interact to mathematical equations like Newton's laws can be used to predict what will happen in an astounding range of settings, including many we have never previously observed \footnote{Although Newton's laws are not fully general, as they break down for extremely massive objects}.

\citet{Gelman2010,gelman2013ask} make a distinction between forward causal inference, the types of \quotes{What if} questions I focus on in this thesis, and reverse causal questions, asking why something occurs. The former aims to identify the effect of a known cause. The latter can be viewed as identifying causes of an effect. They regard forward causal inference as well defined within the counterfactual and graphical model frameworks for causal inference, which we describe in chapter \ref{chap:causal_models}. However, state that \quotes{A reverse causal question does not in general have a well-defined answer, even in a setting where all possible data are made available.} I view this as overly pessimistic, depending on how \quotes{all possible data} is defined. The goal of identifying the causes of an effect can be formalised within the graphical causal model framework. Solving this problem is certainly much more challenging than identifying the effect of a specific intervention on given outcome, since it requires us to test or infer the effect of interventions on many different variables. These practical difficulties may well be overwhelming, particularly in fields such as social science and economics where data sets are relatively small, systems are complex, the variables are difficult to directly manipulate and even relatively simple "What if" questions are hard to resolve conclusively. However, this does not mean that the problem of identifying causes of effects is ill-posed in principle. It can be viewed as a form of causal discovery: the attempt to learn the structure of the causal relationships between variables, on which there is a rich literature which we review briefly in section \ref{sec:causal-discovery}.
\todo{reference a review paper on causal discovery here}

There has traditionally been a large gap between researches in machine learning who focus on prediction, use largely non-interpretable models and researches in statistics, social science and economics who (at least implicitly) aim to answer causal questions and tend to use heavily theory driven models. However, there is relatively little awareness, particularly within the machine learning and data science communities, of what constitutes a causal problem and the implications of for the training and evaluation models. In the next section we emphasise the subtly that can exist in determining if a problem is causal by examining some typical examples.

@@

Consider speech recognition. You say something, which causes sound waves, which are converted converted to a digital signal that Siri maps to words. Whatever action Siri takes is unlikely to change the distribution of words you use, and even less likely to change the function that maps sound waves to text (unless she sends you a DVD on elocution). A similar argument could be made for many applications of machine translation and image classification.

@@

What about forecasting the weather? If you are using a short term forecast to decide whether to pack an umbrella  causality can be ignored - your decision will not effect if it actually rains. However, longer term climate forecasts might (theoretically) lead us to take action on emissions which would then change the weather system. For this we need a (causal) model that allows us to predict the outcome under various different interventions.

@@

This example also highlights a connection between causal models and transparency in machine learning. If we are using a non-causal model to make decision effecting people, there will be a trade-off between the performance and transparency of the model; not because the requirement for transparency restricts us to simple models but because revealing how the model works allows people to change their behaviour to game it.

What about predicting the risk of death in patients with pneumonia? Suppose the goal is to build a model to decide who should be treated in hospital and who can be sent home with antibiotics. If we assume that in hospital treatment is more effective for serious cases, this seems like a straightforward prediction problem. It is not. Depending on how the decision to admit was previously made and what features are included (or omitted) in the model, the relationship between those features and the outcome may change if the model is used to make admission decisions. \citet{Caruana2015} found exactly this effect in a real data set. The model learnt that people suffering asthma were \emph{less} likely to die from pneumonia. They realised this was because doctors were treating such patients very aggressively, thus actually lowering their risk. The issue is not with model, it performed very well at the task for which it was trained, which is to predict who would be likely to die under the original admission and treatment protocols. However, using it to decide how to \emph{change} these protocols could kill. The actual question of interest in this case is what happens to patients with characteristics $X$ if we assign them treatment according to decision rule $Z$.

@@

\todo{is there a mechanical method to go from one framework to another}

\todo{does it matter if variables are discrete or continuous}

@@

A central component of estimating causal effects from observational data is determining if and how we can write expressions for the interventional distributions of interest in terms of observational quantities, which can be measured. We did this on an ad-hoc basis to resolve the examples discussed in section \ref{sec:causal_models}. In this section we describe a principled approach to mapping observational quantities to interventional ones and then, in section \ref{sec:estimation_from_finite_data}, discuss the key issues involved in estimating such expressions from finite sample data. We assume the basic structure of the graph is known. That is, we assume that we can draw a network containing (at a minimum):

@@

\begin{figure}
\centering
\begin{subfigure}[t]{0.48\textwidth}
\centering
\caption{There can be multiple valid adjustment sets. $Z_1$ or $Z_2$ or $\set{Z_1,Z_2}$ all block the backdoor path from $X$ to $Y$.}
\label{fig:adjust_multiple}
\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=1cm,
  thick,lt/.style={latent},ob/.style={observed}]
\node[ob](2){$X$};
\node[ob, right=of 2](3){$Y$};
\node[ob,above=of 3](4){$Z_2$};
\node[ob,above=of 2](5){$Z_1$};
 \path[every node/.style={font=\sffamily\small}]
    (2) edge (3)
    (5) edge (2) edge (4)
    (4) edge (3);
\end{tikzpicture} 
\end{subfigure}
\begin{subfigure}[t]{0.48\textwidth}
\centering
\caption{Conditioning on $Z$ opens the backdoor path $X-U_1-Z-U_2-Y$}
\label{fig:m-graph}
\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=1cm,
  thick,lt/.style={latent},ob/.style={observed}]
\node[ob](1){$Z$};
\node[ob, below left=of 1](2){$X$};
\node[ob, below right=of 1](3){$Y$};
\node[lt,above=of 3](4){$U_3$};
\node[lt,above=of 2](5){$U_1$};
 \path[every node/.style={font=\sffamily\small}]
    (2) edge (3)
    (4) edge (1) edge (3)
    (5) edge (2) edge (1);
\end{tikzpicture} 
\end{subfigure}
\caption{}
\label{fig:adjustment}
\end{figure}

\todo{fix this caption}

@@

This problem differs from the contextual bandit problem in that the learner is not interacting with the environment. As a result, there is no exploration-exploitation trade-off to be made. However, the problem does not reduce to supervised learning, because the label, $y$ is not the desired . In addition, if $\pi$ is allowed to depend $h$ then the samples are not i.i.d. The majority of the literature considers the case where the original policy $\pi$ was stationary ($\pi(\vec{x}_t,h_t) = \pi(\vec{x})$). \citet{Langford2008a} do allow the original policy to be adaptive and prove a high probability bound on the accuracy of their estimator for $\pi'$, albeit with the strong assumption that the original estimator $\pi$ did not depend on $\vec{X}$.

@@

\begin{figure}
\centering
\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=1cm,
  thick,main node/.style={observed},lt/.style={latent}]
\node[main node](1){$\vec{X}$};
\node[main node, below right=of 1](2){$Y$};
\node[main node,  below left=of 1](3){$A$};
 \path[every node/.style={font=\sffamily\small}]
   (3) edge (2)
   (1) edge (2) edge (3);
\end{tikzpicture} 
\caption{Causal graphical model for learning from logged feedback data under the assumption the original policy $\pi$ for selecting actions was stationary and dependent only on some observed context $\vec{X}$, $a \sim \pi(\vec{x})$}
\label{fig:off-policy-evaluation}
\end{figure}

@@

Causal questions are so important

Summarise the amazing stuff I have done

Many interesting future directions for research

>>>>

This thesis is about machine learning and statistical approaches to decision making. How can we learn from data to anticipate the consequence of, and optimally select, interventions or actions? Problems such as deciding which medication to prescribe to patients, who should be released on bail, and how much to charge for insurance are ubiquitous, and have far reaching impacts on our lives. There are two fundamental approaches to learning how to act: reinforcement learning, in which an agent directly intervenes in a system and learns from the outcome, and observational causal inference, whereby we seek to infer the outcome of an intervention from observing the system.

The goal of this thesis to connect and unify these key approaches. I introduce causal bandit problems: a framework that combines causal graphical models, which were developed for observational causal inference, with multi-armed bandit problems, which are a subset of reinforcement learning problems that are simple enough to admit formal analysis. I show that knowledge of the causal structure allows us to transfer information learned about the outcome of one action to predict the outcome of an alternate action, yielding a novel form of structure between bandit arms that cannot be expoited by existing algorithms. I propose an algorithm for causal bandit problems and prove bounds on the simple regret demonstrating it is close to mini-max optimal and strictly better than algorithms that do not use the additional causal information.

@@

\section{Motivation}
Many of the most important questions in science, commerce and our personal lives are about the outcomes of doing something. Will asking people to pay upfront at the doctors reduce long term health expenditure? If we developed a drug to suppress particular genes, could we cure multiple sclerosis and would delaying teen-aged pregnancies improve the outcome for their children.

These are hard questions because they require more than identifying a pattern in data. Correlation is not causation. Causal inference has proven so difficult that there is barely any consensus on even enduring questions like the returns to education or the long-term consequences of early life events – like teenage pregnancy, despite the fact that the variables involved are susceptible to human intuition and understanding.

We now live in a world of data. Hours of our lives are spent online, where every click can be recorded, tiny computers and sensors are cheap enough to incorporate into everything and where the US Institute of Health is considering if all infants should be genetically sequenced at birth. Such data gives us a window into many aspects of our lives at an unprecedented scale and detail but it is messy, complicated and often generated as a by-product of some other purpose. It does not come from the controlled world of a randomised experiment.

The rise of big data sets and powerful computers has seen an explosion in the application of machine learning. From health care, to entertainment and self-driving cars, machine learning algorithms will transform many industries. It has been suggested that the impressive ability of statistical machine learning to detect complex patterns in huge data sets heralds the end of theory \citep{anderson2008end} and that we may be only a short step from \quotes{The Singularity}, where artificial intelligence exceeds our own and then grows exponentially.

However, despite the huge advances in machine learning (in particular deep learning), machine learning algorithms are effective only within narrow problem settings. Getting them to generalise to even slightly different problems or data sets remains very challenging. Deciding how we should act or what policies we should implement requires predictions about how a system will behave if we change it. The correlations detected by standard machine learning algorithms do not enable us to do this, no matter how many petabytes of data they are based on. As machine learning algorithms are incorporated into more and more of the decision making processes that shape the world we live in, it is critical to ensure we understand the distinction between causality and prediction and that we develop techniques for learning how to act that are as effective as those we have for pattern recognition.

@@

The notion of causality has been widely debated in science and philosophy \citep{hume1741treatise,mill1893system,pearson1911grammar,Pearl2000,Lewis2000,woodward2005making,heckman20051,cartwright2007hunting} but is still widely viewed as poorly defined. This has led to a reluctance among applied researchers in many fields to make concrete claims about causality in their work, leading them instead to report that variables are \textit{related}, \textit{correlated} or \textit{associated}. However, the magnitude, direction and even existence of an association depends which variables are controlled for (or included in a regression). Avoiding formalising causation, which is the real question of interest, requires the reader to determine via \quotes{common sense} the implications of the reported associations.

There are two ways in which an association detected in a data set may be non-causal. The first is that the variables concerned may not be related at all, and the association has arisen by chance in that data sample. Given a finite data on enough variables, there is a high probability of finding some that appear correlated even though they are completely unrelated. For example, based on data from the years 1999 to 2009, the age of Miss America is strongly correlated with the number of murders  (in the US) by steam, hot vapours and hot objects \citep{Vigen2015}. We would not expect this relationship to hold in a new sample of data. This form of spurious correlation also has serious repercussions. It lies at the heart of major problems with the process of scientific research: researchers are incentivised to detect effects and thus to explore many possible paths in the process of analysing data and studies that fail to find an effect are less likely to be published. Consequently, the likelihood that reported effects have arisen by chance is underestimated, leading to the conclusion that \quotes{most published scientific results are false} \citep{Ioannidis2005}. This issue is also highlighted by recent crises in replication \citep{OpenScienceCollaboration2015}. This issue can be ameliorated by obtaining more data and by separating learning models from evaluating their performance, for example by evaluating models on a strict hold-out set or on the extent to which their results can be replicated.

However, a strong association, observed in multiple independent studies may still not be causal. The correlation can arise because both variables are consequences of some other, unmeasured factor. For example, the reading ability of children under twelve is strongly correlated with their height, because older children are taller and can read better. However height is not a cause of reading ability because interventions to increase a child's height, for example by giving them growth hormones, would not be expected to improve their reading. Similarly, extra lessons in reading will not make a child grow taller. This problem is fundamentally different to the issue of spurious correlations arising by chance in finite data sets. Obtaining more (even infinitely many more) samples without directly intervening in the system to manipulate the variables does not allow us to separate causation from correlation.

The key distinction between a real, but non-causal, association and a causal relationship is in what happens if we intervene in the system and change one of the variables. In this thesis, I take an interventionist viewpoint of causality: any model or approach designed to predict the outcome of intervening in a system is causal. This viewpoint captures the types of questions that motivate this thesis. How can we change the way we do things to obtain better outcomes?

Causality is often liked to explanation; understanding how and why things happen. I view explanation in terms of compression and generalisabilty: the amount of information about the world a model can capture. This creates a hierarchy in the degree to which models are explanatory, rather than a simple binary distinction. A standard predictive model encodes all the information needed to predict some output given inputs provided the system generating the data does not change. A high-level causal model might be able to predict the outcome of a specific intervention holding all else fixed. More detailed causal models could predict the outcome for a wide range of combinations of interventions conditional on a range of contexts. By considering conditional interventions within our definition of causal questions we also capture mediation: the study of pathways through which one variable causes another \cite{vanderweele2015explanation}. Finally, a model that can distil how elements interact into mathematical equations like Newton's laws can be used to predict what will happen in an astounding range of settings, including many never previously observed \footnote{Although Newton's laws are not fully general, as they break down for extremely massive objects}.

\citet{Gelman2010,gelman2013ask} make a distinction between forward causal inference, the types of \quotes{what if} questions I focus on in this thesis, and reverse causal questions, asking why something occurs. The former aims to identify the effect of a known cause. The latter can be viewed as identifying causes of an effect. They regard forward causal inference as well defined within the counterfactual and graphical model frameworks for causal inference, that we describe in section \ref{sec:causal_models}. However, they state that \quotes{A reverse causal question does not in general have a well-defined answer, even in a setting where all possible data are made available.} I view this as overly pessimistic, depending on how \quotes{all possible data} is defined. The goal of identifying the causes of an effect can be formalised within the graphical causal model framework. Solving this problem is certainly much more challenging than identifying the effect of a specific intervention on a given outcome, since it requires us to test or infer the effect of interventions on many different variables. These practical difficulties may well be overwhelming, particularly in fields such as social science and economics where data sets are often relatively small, systems are complex, the variables are difficult to directly manipulate and even relatively simple "what if" questions are hard to resolve conclusively. However, this does not mean that the problem of identifying causes of effects is ill-posed in principle. It can be viewed as a form of causal discovery: the attempt to learn the structure of the causal relationships between variables, on which there is a rich literature, see for example \citep{}.
\todo{reference a review paper on causal discovery here}

There has traditionally been a large gap between researchers in machine learning who focus on prediction, using largely non-interpretable models and researchers in statistics, social science and economics who (at least implicitly) aim to answer causal questions and tend to use highly theory-driven models. However, there is relatively little awareness, particularly within the machine learning and data science communities, of what constitutes a causal problem and the implications of this for the training and evaluation models. In the next section we emphasise the subtley that can exist in determining if a problem is causal by examining some typical examples.

@@

Consider speech recognition. You say something, which causes sound waves, which are converted to a digital signal that Siri maps to words. Whatever action Siri takes is unlikely to change the distribution of words you use, and even less likely to change the function that maps sound waves to text (unless she sends you a DVD on elocution). A similar argument could be made for many applications of machine translation and image classification.

@@

What about forecasting the weather? If you are using a short term forecast to decide whether to pack an umbrella, causality can be ignored. Your decision will not affect if it actually rains. However, longer term climate forecasts might (theoretically) lead to action on emissions which would then change the weather system. For this we need a (causal) model that allows us to predict the outcome under various different interventions.

@@

This example also highlights a connection between causal models and transparency in machine learning. If we are using a non-causal model to make decisions affecting people, there will be a trade-off between the performance and transparency of the model; not because the requirement for transparency restricts us to simple models, but because revealing how the model works allows people to change their behaviour to game it.

What about predicting the risk of death in patients with pneumonia? Suppose the goal is to build a model to decide who should be treated in hospital and who can be sent home with antibiotics. If we assume that in hospital treatment is more effective for serious cases, this appears to be straightforward prediction. It is not. Depending on how the decision to admit was previously made and what features are included (or omitted) in the model, the relationship between those features and the outcome may change if the model is used to make admission decisions. \citet{Caruana2015} found exactly this effect in a real data set. The model learned that people suffering asthma were \emph{less} likely to die from pneumonia. This was because doctors treated such patients very aggressively, thus actually lowering their risk. The issue is not with model; it performed very well at the task for which it was trained, which is to predict who would be likely to die under the original admission and treatment protocols. However, using it to decide how to \emph{change} these protocols could kill. The actual question of interest in this case is what happens to patients with characteristics $X$ they are assigned treatment according to decision rule (or policy) $\pi(X)$.

@@

Clarify Cheng's Questions - is there a mechanical method to go from one framework to another, does it matter if variables are discrete or continuous. \todo{}

@@

A central component of estimating causal effects from observational data is determining if and how we can write expressions for the interventional distributions of interest in terms of observational quantities, which can be measured. We did this on an ad-hoc basis to resolve the examples discussed in section \ref{sec:causal_models}. In this section we describe a principled approach to mapping observational quantities to interventional ones and then, in section \ref{sec:causal_effect_under_ignorability}, discuss the key issues involved in estimating such expressions from finite sample data. We assume the basic structure of the graph is known. That is, we assume that we can draw a network containing (at a minimum):

@@

\begin{figure}
\centering
\begin{subfigure}[t]{0.48\textwidth}
\centering
\caption{Multiple valid adjustment sets.}
\label{fig:adjust_multiple}
\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=1cm,
  thick,lt/.style={latent},ob/.style={observed}]
\node[ob](2){$X$};
\node[ob, right=of 2](3){$Y$};
\node[ob,above=of 3](4){$Z_2$};
\node[ob,above=of 2](5){$Z_1$};
 \path[every node/.style={font=\sffamily\small}]
    (2) edge (3)
    (5) edge (2) edge (4)
    (4) edge (3);
\end{tikzpicture} 
\end{subfigure}
\begin{subfigure}[t]{0.48\textwidth}
\centering
\caption{Don't condition on all observable variables}
\label{fig:m-graph}
\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=1cm,
  thick,lt/.style={latent},ob/.style={observed}]
\node[ob](1){$Z$};
\node[ob, below left=of 1](2){$X$};
\node[ob, below right=of 1](3){$Y$};
\node[lt,above=of 3](4){$U_3$};
\node[lt,above=of 2](5){$U_1$};
 \path[every node/.style={font=\sffamily\small}]
    (2) edge (3)
    (4) edge (1) edge (3)
    (5) edge (2) edge (1);
\end{tikzpicture} 
\end{subfigure}
\caption{Identifying an optimal adjustment set is not always intuatively obvious. There may be multiple valid adjustment sets. In sub-figure (a) $Z_1$ or $Z_2$ or $\set{Z_1,Z_2}$ all block the backdoor path from $X$ to $Y$. Adding additional variables (even pre-treatment variables) to a valid adjustment set can result in an invalid set as in sub-figure (b). In this case the empty set is a valid adjustment set for the causal effect of $X$ on $Y$, however adding $Z$ would open a backdoor path $X-U_1-Z-U_2-Y$}
\label{fig:adjustment}
\end{figure}

@@

This problem differs from the contextual bandit problem in that the learner is not interacting with the environment. As a result, there is no exploration-exploitation trade-off to be made. However, the problem does not reduce to supervised learning, because of the bandit nature of the feedback - only the reward of the selected action is observed. In addition, if $\pi$ is allowed to depend $h$ then the samples are not i.i.d. The majority of the literature considers the case where the original policy $\pi$ was stationary ($\pi(\vec{x}_t,h_t) = \pi(\vec{x})$). \citet{Langford2008a} do allow the original policy to be adaptive and prove a high probability bound on the accuracy of their estimator for $\pi'$, albeit with the strong assumption that the original estimator $\pi$ did not depend on $\vec{X}$.

@@

\begin{figure}
\centering
\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=1cm,
  thick,main node/.style={observed},lt/.style={latent}]
\node[main node](1){$\vec{X}$};
\node[main node, below right=of 1](2){$Y$};
\node[main node,  below left=of 1](3){$A$};
\node[lt, above=of 2](4){$U$};
 \path[every node/.style={font=\sffamily\small}]
   (3) edge (2)
   (1) edge (2) edge (3)
   (4) edge (1) edge (2);
\end{tikzpicture} 
\caption{Causal graphical model for learning from logged feedback data under the assumption the original policy $\pi$ for selecting actions was stationary and dependent only on some observed context $\vec{X}$, $a \sim \pi(\vec{x})$. The outcome $Y$ may depend on $\vec{X}$ directly, indirectly through a latent variable $U$ or both.}
\label{fig:off-policy-evaluation}
\end{figure}

@@

TODO - write conclusion \todo{}

Here is the problem and why its important

Here is how what I have done contributes to solving thes problems

What remains to be done - directions for future work

- Relaxing assumption that interventional distributions known
- Algorithms for causal bandits in which both contextual and post-action feedback is available.
- Off-policy evaluation for general causal bandits.

Other approaches to identifying causal effects, such instrumental variables, the \quotes{front door} approach could also be applied to off-policy evaluation in some settings.

Ideas developed within the off-policy selection liturature are also relevent classic causal problems within economics etc.
 - optimal action selection (rather than evaluation)
 - quantification of finite time properties of estimators (as opposed to asymptopic efficiency). eg \cite{li2014minimax}

Warm start for (contextual) bandit problem based on potentially biased estimates. (We can detect the presense of a confounder with experimental data).

Alternatives to cross-validation.
